---
title: "Chapter 3:Graph Intro"
output: 
  github_document:
    toc: true
    toc_depth: 2
    fig_width: 8
    fig_height: 6
    dev: jpeg
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = F)
```

# Overview

# Working with graphs

Graphs are built one statement at a time

```{r graph_by_line}
attach(mtcars)
plot(wt, mpg)
abline(lm(mpg~wt))
title("Regression of MPG on Weight")
detach(mtcars)
```

Graphs are saved to disk by wrapping the graph-producing code in a `<file>() ... dev.off()` statement. 

Possible filetypes include:

- `win.metafile()`
- `pdf()`
- `png()`
- `jpeg()`
- `bmp()`
- `tiff()`
- `xfig()`
- `postscript()`

Example code to produce a PDF (not run)

```{r file_out_example, eval=FALSE}
pdf("mygraph.pdf")
  attach(mtcars)
  plot(wt, mpg)
  abline(lm(mpg~wt))
  title("Regression of MPG on Weight")
  detach(mtcars)
dev.off()
```

# Simple example

```{r plot_solo}
dose <- c(20,30,40,45,60)
drugA <- c(16,20,27,40,60)
drugB <- c(15,18,25,31,40)
plot(dose, drugA, type="b")
```

# Graphical parameters

...are commonly called through `par(optionname=value, optionname=value, ...)`

- `pch` specifies the symbol to use for points
- `cex` specifies symbol size as a scale of 1, ex. `5 = scale by 500%`
- `lty` specifies line type
- `lwd` specifies line width as a scale of 1

## Point and line aesthetics

All possible point symbols.

```{r pch_options}
x <- rep(1:5,6)[1:26]
y <- c(rep(6:2,each=5),1)
pch_vals <- 0:25
plot(x-0.25, y,
     pch=pch_vals,
     xlim=c(min(x)-0.25, max(x)+0.25),
     bty = "n",
     axes = F, xlab=NA, ylab=NA)
mtext("pch options",side=3,cex=2,line=1.2,adj=0)
axis(side=2,labels=F,tck=F)
text(x,y,pch_vals)
```

All possible line types.

```{r line_types}
y <- c(1:6)
plot(y,
     lty=y,
     bty="n",
     axes=F, xlab=NA, ylab=NA,
     pch=NA)
mtext("ltw options",side=3,cex=2,line=1.2,adj=0)
text(x=1,y+0.15,rev(y))
abline(h=y,lty=rev(y))
```

## Colors

Colors are specified by index `col=1`, name `col="white"`, hex code `col=#FFFFFF`, RGB `col=rdb(1,1,1)`, or HSV `col=hsv(0,0,1)`. 

Parameters for customizing various plot components:

- `col` set default plotting color. If a vector is provided, the plot will cycle through the vector while producing elements
- `col.axis` sets axis text
- `col.lab` sets axis labels
- `col.main` sets title
- `col.sub` sets subtitle
- `fg` sets plot foreground
- `bg` sets plot background

`colors()` returns all available color names. 

```{r colors}
head(colors())
```

R also has several functions for contiguous colors, including `rainbow()`, `heat.colors()`, `topo.colors()`, and `cm.colors()`.

```{r rainbow_pie}
n <- 30
colors <- rainbow(n)
pie(rep(1, n), labels=colors, col=colors)
mtext("Rainbow", side=3, cex=2, line=1.2, adj=0)
mtext("pretty...",side=3,cex=1,line=-0.25,adj=0)
```

```{r heat_pie_topo}
colors <- heat.colors(n)
pie(rep(1, n), labels=colors, col=colors, main = "Heat")
mtext("Heat",side=3,cex=2,line=1.2,adj=0)
mtext("for heat and other neat things",side=3,cex=1,line=-0.25,adj=0)
```

```{r pie_topo}
colors <- topo.colors(n)
pie(rep(1, n), labels=colors, col=colors)
mtext("Topographic",side=3,cex=2,line=1.2,adj=0)
mtext("for maps",side=3,cex=1,line=-0.25,adj=0)
```

```{r pie_cm}
colors <- cm.colors(n)
pie(rep(1, n), labels=colors, col=colors)
mtext("CM",side=3,cex=2,line=1.2,adj=0)
mtext("what does 'CM' stand for?",side=3,cex=1,line=-0.25,adj=0)
```

## Text

Text size:

- `cex` size relative to 1
- `cex.axis`
- `cex.lab`
- `cex.main`
- `cex.sub`

Font family, point size, style:

- `font` font for plotted text
- `font.axis` axis text
- `font.lab` axis labels
- `font.main` titles
- `font.sub` subtitles
- `ps` font point size text size `=ps*cex`
- `family` font family, standard values are `serif`, `sans`, and `mono`

## Graph and margin dimensions

Dimension and margin size parameters:

- `pin` plot dimensions (width, height) in inches (inches? strange...)
- `mai` margin size, `c(bottom, left, top, right)` in inches
- `mar` margin size, `c(bottom, left, top, right)` expressed in lines

Margin example. Source: https://www.r-bloggers.com/setting-graph-margins-in-r-using-the-par-function-and-lots-of-cow-milk/

```{r margin_example}
plot(1:10,ann=FALSE,type="n",xaxt="n",yaxt="n")
for(j in 1:4) for(i in 0:10) mtext(as.character(i),side=j,line=i)
```

# - - - - - - STOPPED HERE 8/19/16 - - - - - - - 

# Adding text, customized axes, and legends

## Titles

## Axes

## Reference lines

## Legend

## Text annotations

## Math annotations

# Combining graphs

## Creating a figure arrangement with fine control

# Recap

